\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}
\usepackage{textcomp} %\textdegree
%SetFonts
%SetFonts

\usepackage[square]{natbib} % <- new

\title{Research areas and bibliography relevant to the Strait of Magellan}
\author{M. Williams...}
%\date{}							% Activate to display a given date or no date

\begin{document}
\maketitle
%\section{}
%\subsection{}
\subsection{geology, morphology, sediment}

\citep{simeoni97}

\subsection{Hydrodynanamics of the region} 

Early work by University of British Columbia made surveys in the 1970s \citep{pickard71}.

Fronts in the northern fjords \citep{calvete11}. Part of a Continental Shelf Research special issue on the Patagonian fjords \cite{pantoja11}.

Glacier melting and rivers on nutrient \citep{vargas18}.

On the bio-optics \citep{lutz16}.

\subsection{Environmental/aquaculture}

\citep{niklitschek13} cites \citet{iriarte10} as Patagonia having the largest estuary on earth. The \citet{niklitschek13} review is on aquiculture pressures in Southern Chile, and urges no further contracts while appropriate management is decided. Until 2008, 70 - 90\% of Chilean salmon harvest was in the Chilo{\'e} inner sea, it then expanded rapidly to  Ays{\'e}n with an infection salmon anaemia virus (ISAV) crisie between 2008-10. Development now (2013) more in Ays{\'e}n to keep density down (a contributing factor in ISAV) and empirical evidence that sea lice and ISAV are slower in cooler water (of Ays{\'e}n). Impacts of salmon farming - locally increased sedimentation under cages, benthic species, environments change. Two to five-fold increase in nutrients has been reported, 50\% less species richness. Remediation studies are few and not very long, it takes 5-6 years for the chemistry and physics to return, but more than 7 for the benthic communities - depends strongly on local oceanogrphy. Other impacts: antifouling and antibiotics. Fjords in Patagonia are nutrient limited, oceanic Subantarctic Waters (SAAW) are the main source of nutrients \citep{iriarte10}, but despite incredible release of N and P from fish farms, little evidence that they are changing the nearby pelagic environments. 

One of the largest oil spills in history, 1974 VLCC Metula oil spill on the northern side of the Primera Angostura in the Magellan Strait. 

\citep{iriarte18}

\subsection*{Tidal energy in inlets}

Work has been done to study tidal energy potential of Admiralty Inlet, located in the Puget Sound in Washington state, USA. The inlet has currents up to (I think) 3 m/s on spring tides. Marine energy related work: \citep{polagye13, kawasethyng10, thomson12, polagye09}. 

There is also extensive study of the area in terms of fjord dynamics, form drag, estuarine physics \citep[e.g.][]{warner13, geyer82}.


Tides in the Magellan Strait: \citep{conteras04, medeiros88}.

\subsection*{Oceanography of the Patagonian shelf, influence of Magellan inputs}
Productivity on the Patagonian shelf \citep{song16}. Numerical modeling on the Patagonian shelf is mostly based on the work of V. Combes \citep{combes14}. The Magellan plume \citep{palma12} which cites the work by Panella et al. 1991 as model inputs.

And some very recent work on how the Magellan strait influnces dynamics on the shelf \citep{guihou20, brun20}


\subsection*{Prior measurements in the Strait of Magellen}
Measurements that have been made in the Magallanes... \citep{antezana99}

Papers worth reading that I can't (yet) find a digital copy: 

\begin{itemize}
\item Panella et al. 1991 A preliminary contribution to understanding the hydrological characteristics of the Strait of Magellan: Austral Spring 1989, S Panella, A Michelato, R Perdicaro, G Magazzu? - \textit{Bollettino Di Oceanologia Teorica Ed Applicata}, 1991

\item Michelato A, Accerboni E, Berger P (1991) Current meter observations in the eastern and central sectors of the Strait of Magellan \textit{Boll Oceanol Teor Appl} 9(2-3):261-271
\end{itemize}

Measurements near a glacial fjord on the MS \citep{ValleLevinson2006}


\subsection*{non-hydrodynamic focus in magellan strait}

\citep{talbot1997meps}

\subsection{patagonia in general}

More north: \citep{narvaez19}.

\subsection{Otway, Skyring, Fitzroy}

The Senos are there because they used to be the extent of ice lobes \citep{darvill17}, possibly rapid formation \citep{lovell12}.

Papers with any reference to water quality/physical variables in these senos: \citep{mansilla13, rodriguez18, valdenegro03, rios16, mansilla13_AIP}. The \citet{rios16} paper has salinity and temperature observations in Otway, and shows that there are thin-layers of Chl-A related to water column stratification and mixing. \citep{cuevas19} also has measurments in Otway. There generally seem to be more Otway measurements than Skyring. 

There is a nice bathymetry map of Seno Skyring in \citep{breuer13}.

\citep{kilian07} shows salinity in Skyring between 17 and 19 ppt. 

\bibliography{magallanes}{}
\bibliographystyle{apalike}

\end{document}  